Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-23 paper/v2/paper-edited.tex
final edits

Line 2 of this policy
restricts traffic destined to known customer prefixes (\CD{PCust}) to only follow paths that end at the customer. In addition, it enforces the network-wide preference that traffic leaves through \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}.
%defines the paths that exit our network to customer prefixes---\CD{ExitCust} defines paths through \CD{R1} and \CD{R2}, and in the event
%that connections to the customer through both routers fail,
%a backup route (\CD{\End(Cust)}) is defined that admits traffic along
%any path (including those through other networks) that ends at the customer.
%Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.
Line 3 applies to any other traffic not matching \CD{PCust} and allows the traffic to leave through any direct neighbor with the usual preferences of \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}. To summarize our progress, the \CD{Routing} policy
implements P1, P3, and P4 from Figure~\ref{fig:example1}.

@@

\begin{code}
\Define NoTrans =
  \{\True \Path \NOT{}transit(Peer,Prov) \AND
           \NOT{}cust-transit(Cust,Prov)\}
\end{code}
\noindent
The \CD{NoTrans} constraint requires that all traffic not follow a path that transits the network between \CD{Peer} and \CD{Prov}.
Additionally, it prevents traffic from ever following paths that leave our network and later go through both \CD{Prov} and \CD{Cust}.  To implement both \CD{Routing}
and \CD{NoTrans} simultaneously, we simply conjoin them: \CD{Routing \AND{} NoTrans}.
%
%% \begin{lstlisting}[mathescape=true]
%% Routing $\AND$  NoTransit
%% \end{lstlisting}
%

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints. The first kind of constraint is a test on the type of route and a corresponding set of preferred regular paths or a control constraint. Regular paths are regular expressions where the base characters are abstract locations representing either a router or an external AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expression abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisements with a particular community value (\IE, an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The next step rewrites the high-level constraints such as $\Enter$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $r \cdot (s \Prefer t) \cdot u$ is lifted to $(r \cdot s \cdot u) \Prefer (r \cdot t \cdot u)$ by distributing the preference over the sequence operator. In general, we employ the following distributivity equivalences:
%
\[
\begin{array}{c}
  r \odot (s_1 \Prefer \dots \Prefer s_n) = (r \odot s_1) \Prefer \dots \Prefer (r \odot s_n) \\
  (s_1 \Prefer \dots \Prefer s_n) \odot r = (s_1 \odot r) \Prefer \dots \Prefer (s_n \odot r)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $r$ is a policy with a single preference.
%
In cases where $r$ does not contain a single preference, such as $(s ~\Prefer~ t) \cdot (u ~\Prefer~ v)$, it is not clear which of the paths $s \cdot v$ or $t \cdot u$ is preferred. \sysname rejects such ambiguous policies, requiring instead that operators explicitly specify which paths to prefer --- for example as $(s \cdot u) ~\Prefer~ (s \cdot v) ~\Prefer~ (t \cdot u) ~\Prefer~ (t \cdot v)$.
%
Policies that contain preferences nested under a unary operator (\IE, \textit{star} or \textit{negation}) are also rejected by \sysname as invalid.

@@

Intuitively, the PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$ and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $n$ such that the preference function $P(n) = \set{i_1, \dots, i_k}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_k$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possibility is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton cannot accept the current path or any extension of it. Since
node $(W,5,-)$ is in an accepting state for the first automaton, it indicates that this path has preference 1.

@@

To demonstrate the difficulty of generating device-local preferences, consider the simple policy for the topology in Figure~\ref{fig:unimplementable}, which says to prefer the top path over the bottom path:
%
$(\CD{A} \cdot \CD{B} \cdot \CD{D} \cdot \CD{E} \cdot \CD{G})  \Prefer (\CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \CD{F} \cdot \CD{G})$.
%
How could such a policy be implemented in BGP? Suppose we set the local preferences to have $D$ prefer $E$ over $F$, and have $A$ prefer $B$ over $C$. This works as expected under normal conditions, however, if the A--B link fails, then suddenly $D$ has made the wrong decision by preferring $E$. Traffic will now follow the
$A \cdot C \cdot D \cdot E \cdot G$ path, even though this path was not allowed by the policy. Thus, the distributed implementation has used a route that is not allowed by the policy. To make matters worse, the second preference for the path $A \cdot C \cdot D \cdot F \cdot G$ is available in the network but not being used. Thus, a path for the best possible route available exists in the network but is not being used by the distributed implementation.
%
The first problem could be fixed by tagging and filtering route advertisements appropriately so that $C$ rejects routes that go through $E$, however the second problem cannot be fixed. In fact, this policy cannot be implemented in BGP in a way that is policy compliant under all failures since $D$ cannot safely choose between $E$ and $F$ without knowing whether the A--B link is available.

@@

The problem of determining local preferences for each router is reflected in the structure of the PGIR. Whenever a given router appears as multiple shadow nodes in the PGIR, the compiler must decide which shadow to prefer.
%
In the example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in shadow $(C,-,2)$ or from $D$ in shadow $(C,3,2)$.
%
%
The compiler must determine a \emph{total ordering} of shadow nodes for each router, which reflects the relative preference of advertisements received in each shadow and should be consistent with path ranks in routing policy.
%
For example, if $C$'s shadow $(C,3,2)$ can be preferred to $(C,-,2)$, written as $(C,3,2) \leq_{lp} (C,-,2)$, $C$ can prefer advertisements from $(D,2,2)$ over $(E,-,2)$. $D$ and $E$ tag their advertisements to let $C$ know which shadow sent the advertisement. \$\ref{sec:abgp} discusses tagging in detail.

@@

For example, recall the datacenter example from \S\ref{sec:motivation}, with the policy
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the PGIR for \CD{PG1}. Since we know aggregation will occur at $X$, and that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first and would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is now disconnected from $X$, 2 is a lower bound on the number of failures required to introduce an aggregation black hole for prefix \CD{PG1}. This process is repeated for other aggregation locations (\EG, $Y$).

@@

Our \sysname compiler is implemented in roughly 6700 lines of F\# code. It includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, the no-export community, as well as for ensuring at least k-failure safety for aggregate prefixes. Since each prefix has a separate routing policy, we compile each routing policy in parallel. Currently, \sysname supports generating Quagga router configurations out of the box. Users can add new vendor-specific adapters to translate from ABGP to other router configuration languages, or incorporate the compiler into an existing template-based system, \EG, by mixing the \sysname-generated BGP configuration with other, non-BGP configuration elements.

>>>>

Line 2 of this policy
restricts traffic destined to known customer prefixes (\CD{PCust}) to only follow paths that end at the customer. In addition, it enforces the network-wide preference that traffic leaves through \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}.
%defines the paths that exit our network to customer prefixes---\CD{ExitCust} defines paths through \CD{R1} and \CD{R2}, and in the event
%that connections to the customer through both routers fail,
%a backup route (\CD{\End(Cust)}) is defined that admits traffic along
%any path (including those through other networks) that ends at the customer.
%Line 3 states that traffic for internal prefixes must end in our network, and is otherwise unconstrained.
Line 3 applies to any other traffic not matching \CD{PCust} and allows the traffic to leave through any direct neighbor with the usual preference of \CD{R1} over \CD{R2} over \CD{Peer} over \CD{Prov}. To summarize our progress, the \CD{Routing} policy
implements P1, P3, and P4 from Figure~\ref{fig:example1}.

@@

\begin{code}
\Define NoTrans =
  \{\True \Path \NOT{}transit(Peer,Prov) \AND
           \NOT{}cust-transit(Cust,Prov)\}
\end{code}
\noindent
The \CD{NoTrans} constraint requires that all traffic not follow a path that transits our network between \CD{Peer} and \CD{Prov}.
Additionally, it prevents traffic from ever following paths that leave our network and later go through both \CD{Prov} and \CD{Cust}.  To implement both \CD{Routing}
and \CD{NoTrans} simultaneously, we simply conjoin them: \CD{Routing \AND{} NoTrans}.
%
%% \begin{lstlisting}[mathescape=true]
%% Routing $\AND$  NoTransit
%% \end{lstlisting}
%

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints. The first kind of constraint is a test on the type of route and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations representing either a router or an external AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any location. We also use the standard regular expression abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisements with a particular community value attached (\IE, an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The next step rewrites the high-level constraints such as $\Exit$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $r \cdot (s ~\Prefer~ t) \cdot u$ is lifted to $(r \cdot s \cdot u) \Prefer (r \cdot t \cdot u)$ by distributing the preference over the sequence operator. In general, we employ the following distributivity equivalences:
%
\[
\begin{array}{c}
  r \odot (s_1 \Prefer \dots \Prefer s_n) = (r \odot s_1) \Prefer \dots \Prefer (r \odot s_n) \\
  (s_1 \Prefer \dots \Prefer s_n) \odot r = (s_1 \odot r) \Prefer \dots \Prefer (s_n \odot r)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $r$ is a policy with a single preference.
%
In cases where $r$ does not contain a single preference, such as $(s ~\Prefer~ t) \cdot (u ~\Prefer~ v)$, it is not clear which of the paths $s \cdot v$ or $t \cdot u$ is preferred. \sysname rejects such ambiguous policies, requiring instead that operators explicitly specify which paths to prefer --- for example as $(s \cdot u) ~\Prefer~ (s \cdot v) ~\Prefer~ (t \cdot u) ~\Prefer~ (t \cdot v)$.
%
Policies that contain preferences nested under a unary operator (\IE, \textit{star} or \textit{negation}) are also rejected by \sysname as invalid.

@@

Intuitively, the PGIR tracks the policy states of each automaton as route announcements move between locations.
%
Consider the topology in Figure~\ref{fig:example-compilation}. Suppose we want a primary route from neighbor W that allows traffic to enter the network at $A$ and utilize the C--D link before leaving the network (through $X$ or $Y$). As a backup, we also want to allow traffic to enter the network from $B$, in which case the traffic can also utilize the C--E link before leaving the network. For simplicity, we assume that the route ends in either $X$, $Y$, or $Z$. The RIR for the policy could be written as:
%
$$(\CD{W} \cdot \CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \Out) \Prefer (\CD{W} \cdot \CD{B} \cdot \In^+ \cdot \Out)$$
%
Figure~\ref{fig:example-compilation} shows the policy automata for each regular expression preference. Since we are interested in the flow of control messages, the automata match backwards.
%
The figure also shows the PGIR after intersecting the topology and policy automata. Every path in the PGIR corresponds to a concrete path in the topology. In particular, every path through the PGIR that ends at a node $n$ such that the preference function $P(n) = \set{i_1, \dots, i_k}$ is non-empty, is a valid topological path that satisfies the policy constraints and results in a particular path with preferences $i_1$ through $i_k$.
%
For example, the path $X \cdot D \cdot C \cdot A \cdot W$ is a valid path in the topology that BGP route announcements might take, which would lead to obtaining a path with the lowest (best) rank of $1$.
BGP control messages can start from peer X, which would match the $\Out$ transition from both automata, leading to state $1$ in the first automaton, and state $1$ in the second automaton. This possibility is reflected in the product graph by the node with state $(X,1,1)$. From here, if X were to advertise this route to D, it would result in the path $D \cdot X$, which would lead to state $2$ in the first automaton, and state $2$ in the second automaton, and so on.
%
The ``--" state indicates the corresponding automaton cannot accept the current path or any extension of it. Since
node $(W,5,-)$ is in an accepting state for the first automaton, it indicates that this path has rank 1.

@@

To demonstrate the difficulty of generating device-local preferences, consider the simple policy for the topology in Figure~\ref{fig:unimplementable}, which says to prefer the top path over the bottom path:
%
$(\CD{A} \cdot \CD{B} \cdot \CD{D} \cdot \CD{E} \cdot \CD{G})  \Prefer (\CD{A} \cdot \CD{C} \cdot \CD{D} \cdot \CD{F} \cdot \CD{G})$.
%
How could such a policy be implemented in BGP? Suppose we set the local preferences to have $D$ prefer $E$ over $F$, and have $A$ prefer $B$ over $C$. This works as expected under normal conditions, however, if the A--B link fails, then suddenly $D$ has made the wrong decision by preferring $E$. Traffic will now follow the
$A \cdot C \cdot D \cdot E \cdot G$ path, even though this path was not allowed by the policy. Thus, the distributed implementation has used a route that is not allowed by the policy. To make matters worse, the second preference for the path $A \cdot C \cdot D \cdot F \cdot G$ is available in the network but not being used. Thus, a path for the best possible route available after the A--B failure exists in the network, but the distributed implementation will not find it.
%
The first problem could be fixed by tagging and filtering route advertisements appropriately so that $C$ rejects routes that go through $E$, however the second problem cannot be fixed. In fact, this policy cannot be implemented in BGP in a way that is policy compliant under all failures since $D$ cannot safely choose between $E$ and $F$ without knowing whether the A--B link is available.

@@

The problem of determining local preferences for each router is reflected in the structure of the PGIR. Whenever a given router appears as multiple shadow nodes in the PGIR, the compiler must decide which shadow to prefer.
%
In the example from Figure~\ref{fig:example-compilation}, the topology node $C$ can receive an advertisement from $E$ in shadow $(C,-,2)$ or from $D$ in shadow $(C,3,2)$.
%
%
The compiler must determine a \emph{total ordering} of shadow nodes for each router, which reflects the relative preference of advertisements received in each shadow and should be consistent with path ranks.
%
For example, if $C$'s shadow $(C,3,2)$ can be preferred to $(C,-,2)$, written as $(C,3,2) \leq_{lp} (C,-,2)$, $C$ can prefer advertisements from $(D,2,2)$ over $(E,-,2)$. $D$ and $E$ tag their advertisements to let $C$ know which shadow sent the advertisement. Section \ref{sec:abgp} discusses tagging in detail.

@@

Recall the datacenter example from \S\ref{sec:motivation}, with the policy
\CD{PG1} \Path \text{ }\End(\CD{A}), where \CD{PG1} falls under the \CD{PG} aggregate. Figure~\ref{fig:aggregation-safety} shows the PGIR for \CD{PG1}. Since we know aggregation will occur at $X$, and that the \CD{PG1} prefix will originate at $A$, we can compute the number of failures it would take to disconnect $A$ from $X$. We could remove the $A$--$D$--$X$ path first and would then need to remove any other $A$--$D$ or $D$--$X$ links from the PGIR (in this case none). Next, we could remove the links along the $A$--$C$--$X$ path, repeating the process. Because $A$ is now disconnected from $X$, 2 is a lower bound on the number of failures required to introduce an aggregation black hole for prefix \CD{PG1}. This process is repeated for other aggregation locations (\EG, $Y$).

@@

Our \sysname compiler is implemented in roughly 6700 lines of F\# code. It includes command-line flags for enabling or disabling the use of the BGP MED attribute, AS path prepending, the no-export community, as well as for ensuring at least k-failure safety for aggregate prefixes. Since each predicate has a separate routing policy, we compile each routing policy in parallel. Currently, \sysname supports generating Quagga router configurations out of the box. Users can add new vendor-specific adapters to translate from ABGP to other router configuration languages, or incorporate the compiler into an existing template-based system, \EG, by mixing the \sysname-generated BGP configuration with other, non-BGP configuration elements.
